---
title: "Bitcoin price and UTXO"
author: "Daran Thach"
date: "3/17/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The blockchain universe can be divided into two main architectures for record-keeping of output transactions: 1) UTXO-Based: Bitcoin, Bitcoin Cash, Litecoin, Dash, and 2) Account-Based: Ethereum, Ethereum Classic.  Beyond being a critical architecture of some of the main blockchains in the market, UTXOs are also a unique source of intelligence, and thus may help us understand unique characteristics of buyers and sellers.  Using blockchain.com data, I used R to import, find descriptive stats, cleaned, prepared UTXO/price data of BTC for the last 2 years, in order to compile correlation analysis between the two.

```{r include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

#import the bitcoin excel values
btcdata <- read.csv("btc_utxo_price.csv")

#find descriptive stats
summary(btcdata)
#find variance and price samples
var(btcdata)
head(btcdata$price_usdavg)
#convert factor to date data
as.Date.character(btcdata$endweek, tryFormats = c("%m/%d/%Y"))

# Visualize time series and plots
library(ggplot2)
library(tidyr)
library(dplyr)

# Multiple ts plot change aesthetics and labels
chart <- btcdata %>%
  select(endweek, utxo_average, price_usdavg) %>%
  gather(key = "variable", value = "value", -endweek)
head(chart, 3)
g <- ggplot(chart, aes(x = endweek, y = value)) + 
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(values = c("#33CC33", "#0066FF")) +
  theme_minimal()
g <- g + labs(title = "BTC Unspent Transaction Output", 
  caption = "Source: blockchain.info")
g

# Correlate the ts plots by determining linearity and nml dist
library(ggpubr)
# plot for utxo
ggqqplot(btcdata$utxo_average, ylab = "UTXO")
# plot for price
ggqqplot(btcdata$price_usdavg, ylab = "Price USD")

# Use Spearman rank correlation coefficient
rank <-cor.test(btcdata$utxo_average, btcdata$price_usdavg, method = "spearman")
rank
ggscatter(btcdata, x = "utxo_average", y = "price_usdavg", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "spearman",
          xlab = "Unspent transaction output", ylab = "Price in USD")

```
# End of R script
The above outputs generated time series plots upon changing aesthetics/labels, then applied statistical linearity testing, and Spearman rank correlation.  Revealed UTXO moderately correlated to price of BTC.